### 2022-03-17 satfol/satfol.tex
more

>>>>

\subsection{Refutational Completeness}
\label{ssec:predicate-elimination-refutational-completeness}

\newcommand\FInf{\mathit{FInf}}
\newcommand\FLInf{\mathit{FLInf}}
\newcommand\Red{\mathit{Red}}
\newcommand\RedI{\Red_\mathrm{I}}
\newcommand\RedF{\Red_\mathrm{F}}
\newcommand\LRed{\mathit{LRed}}
\newcommand\LRedI{\LRed_\mathrm{I}}
\newcommand\LRedF{\LRed_\mathrm{F}}
\newcommand\gnd{\mathcalx{G}}
\newcommand\pow[1]{\mathcalx{pe}(#1)}
\newcommand\Lab{\mathbf{L}}
\newcommand\FFLab{{\mathbf{FL}}}
\newcommand\Inv{\mathit{Inv}}

\newcommand\RedWk{\smash{\Red\Wksup}}
\newcommand\RedIWk{\smash{\Red_\mathrm{I}\Wksup}}
\newcommand\RedFWk{\smash{\Red_\mathrm{F}\Wksup}}

\newcommand\Wksym{\flat}
\newcommand\Wksup{^\Wksym}
\newcommand\modelsWk{\models\Wksup}
%\newcommand\modelsWkcapgnd{\models^{\Wksym{}\gnd}_\cap}
\newcommand\modelsWkcapgnd{\modelsWk}
\newcommand\LRedWk{\LRed^{\Wksym}}
\newcommand\LRedWksub{\LRed^{\Wksym,\sqsupset}}
\newcommand\LRedFWksub{\LRedF^{\Wksym,\sqsupset}}
\newcommand\LRedIWk{\LRedI^{\Wksym}}
\newcommand\Wk[1]{#1^\Wksym}
\newcommand\prop{o}
\newcommand\true{\top}
\newcommand\false{\bot}

\looseness=-1
Hidden-literal-based techniques fit within the traditional framework of
saturation, because they delete or reduce a clause based on the \emph{presence}
of other clauses. In contrast, predicate elimination relies on the
\emph{absence} of clauses from the proof state. We can still integrate it with
superposition as follows: At every $k$th iteration of the given clause
procedure, perform predicate elimination on $\mathcal{A}
\cup \mathcal{P}$, and add all new clauses to $\mathcal{P}$.

One may wonder whether such an approach preserves the refutational
completeness of the calculus. The answer is no.
%
\begin{conf}
To see why, consider the following \emph{binary splitting} rule
based on Riazanov and Voronkov \cite{riazanov-voronkov-2001}:
%
\[\namedsimp{BS}{C \cor D}{\cst{p} \cor C \quad D \cor \neglit{\cst{p}}}\]
%
Provisos: $C$ and $D$ have no free variables in common, $\cst{p}$ is fresh, and
$\cst{p}$ is $\prec$-smaller than $C$ and $D$. Since the conclusions are
smaller than the premise, the rule can be applied aggressively as a
simplification. But notice that the effect of splitting can be undone by
singular predicate elimination, possibly giving rise to loops
$\infname{BS}, \infname{SPE}, \infname{BS}, \infname{SPE}, \dotsc$.
This breaks completeness.

Our solution is to curtail the entailment relation used by the redundancy
criterion to disallow splitting-like simplifications. Weak entailment $\modelsWk$
is defined via an ad hoc nonclassical logic so that $\{\cst{p} \cor C{,}\;
\neglit{\cst{p}} \cor C\} \notmodelsWk \{C\}$ and yet $\modelsWk \{\cst{p} \cor
\neglit{\cst{p}}\}$.
%
More precisely, this logic is defined via an
encoding: $M \modelsWk N$ if and only if $\Wk{M} \models \Wk{N}$, where
%
$\Wk{\cst{p}(\tuple{t}{})} = \cst{p}(\tuple{t}{}) \noteq \false$,
$\Wk{\lnot\,\cst{p}(\tuple{t}{})} = \cst{p}(\tuple{t}{}) \noteq \true$, and
$\Wk{L} = L$ otherwise.
%
Moreover, the type $\prop$ may be interpreted as any set of cardinality at least
2, and $\false$ must be a distinguished symbol interpreted differently
from~$\top$.

The standard redundancy criterion $\RedWk$ based on $\modelsWk$ supports all the
familiar deletion and simplification techniques except splitting. Using $\RedWk$
not only prevents looping, but it also enables the use of
the given clause procedure, because any redundant inference
according to $\RedWk$ remains redundant after SPE or DPE. As usual, the devil is
in the details, and the details are in the report \cite{our-report}.
\end{conf}

\begin{rep}
To see why, consider the following \emph{binary splitting} rule
based on Riazanov and Vo\-ro\-nkov~\cite{riazanov-voronkov-2001}:
%
\[\namedsimp{BS}{C \cor D}{\cst{p} \cor C \quad D \cor \neglit{\cst{p}}}\]
%
Provisos: $C$ and $D$ have no free variables in common, $\cst{p}$ is fresh, and
$\cst{p}$ is $\prec$-smaller than $C$ and $D$. Since the conclusions are
smaller than the premise, the rule can be applied aggressively as a
simplification. But notice that the effect of splitting can be undone by
singular predicate elimination, possibly giving rise to loops
$\infname{BS}, \infname{SPE}, \infname{BS}, \infname{SPE}, \dotsc$.
Clearly, we need to curtail predicate elimination.

Under which conditions is predicate elimination refutationally complete? To
answer this question, we employ the saturation framework of Waldmann, Tourret,
Robillard, and Blanchette \cite{wtrb-20-sat-framework}. Let
$(\FInf, \Red)$ be the base calculus without predicate elimination---e.g.,
resolution or superposition inferences together with the standard
redundancy criterion \cite[Sect.~4.2]{bg-01-resolution}. The
inference system $\FInf$ is a set of inferences $(C_n,\ldots,C_1,C_0)$, for $n
\ge 1$, where $C_n,\ldots,C_1$ are the premises and $C_0$ is the conclusion.
$C_1$ is called the main premise. %; $C_n,\ldots,C_2$ are the side premises.
The redundancy criterion is a pair $\Red = (\RedI, \RedF)$ where $\RedI$
determines which inferences can be omitted and $\RedF$ is used to remove
clauses.

Next, consider an abstract proving process working on a single clause set.
Let $\rhd_{\Red}$ denote the transition relation that supports (1)~adding
arbitrary clauses and (2)~removing clauses deemed useless by $\RedF$.
Typically, the added clauses are the result of performing inferences and are
entailed by the premises, but other clauses can be added as well. A
\emph{$\rhd_{\Red}$-derivation} is a finite or infinite sequence of clause sets
$N_0 \rhd_{\Red} N_1 \rhd_{\Red} \cdots$.

We fix a finite set $\mathbf{P}$ of predicate symbols that may be subjected to
predicate elimination. These might include all the predicate symbols occurring
in the input problem, but exclude any symbols introduced by splitting or other
rules. Given a clause or clause set $N$, we write $\mathbf{P}(N)$ to denote the
set of all predicate symbols from $\mathbf{P}$ occurring in $N.$
%
Let $\rhd_\mathbf{P}$ denote the elimination of a
singular or defined predicate symbol from $\mathbf{P}$. A \emph{mixed
derivation} consists of transitions either of the form $N
\rhd_\mathbf{P} N'$ or of the form $N \rhd_{\Red} N'$ where
$\mathbf{P}(N) \supseteq \mathbf{P}(N')$.
Because $\mathbf{P}$ is finite, any mixed derivation consists of at most
finitely many $\rhd_\mathbf{P}$ transitions. Hence, in any derivation,
there exists an index~$k$ from which all transitions are standard
${\rhd}_{\Red}$-transitions.

This suggests the following path to completeness: Pretend that the transitions
between $N_0$ and $N_k$ are merely preprocessing and start the actual
derivation at $N_k$. This works at the abstract level of derivations on single
clause sets. It fails, however, for an actual saturation prover that
distinguishes between passive and active clauses.

\begin{exa}
\label{ex:cex-predelim-sat}
The counterexample below is based on the given clause prover \textsf{GC}
from the saturation framework. It shows how predicate elimination can break
\textsf{GC}'s key invariant, which states that all inferences between active
clauses are redundant. Breaking the invariant means that the limit might be
unsaturated, breaking the refutational completeness proof.

We use superposition with the order $\cst{a} \prec
\cst{b} \prec \cst{c} \prec \cst{d}$ and without selection.
Assume $\cst{a} \in \mathbf{P}$ and
suppose we start with the satisfiable clause set %consisting of
%
%\begin{linenomath*}
\begin{align*}
  & \neglit{\cst{a}} \cor \MAX{\cst{d}}
  && \neglit{\cst{a}} \cor \MAX{\neglit{\cst{d}}}
  && \cst{a} \cor \cst{b} \cor \MAX{\cst{c}}
  && \cst{c} \cor \MAX{\cst{d}}
  && \cst{b} \cor \MAX{\neglit{\cst{d}}}
\end{align*}
%\end{linenomath*}
%
where gray boxes mark maximal (i.e., eligible) literals. Suppose the prover makes
$\cst{c} \cor \cst{d}$ and $\cst{b} \cor \neglit{\cst{d}}$ active. From these
two clauses, a superposition inference~$\iota$ could derive the conclusion
$\cst{b} \cor \cst{c}$. However, the three passive clauses are
$\prec$-smaller than $\iota$'s main premise $\cst{b} \cor \neglit{\cst{d}}$ and
collectively entail $\iota$'s conclusion. This means that $\iota$ is
redundant and can be ignored.

If the prover now eliminates the predicate $\cst{a}$ using \infname{SPE}, the
passive set is reduced to $\{\cst{b} \cor \cst{c} \cor \cst{d}{,}\;
\cst{b} \cor \cst{c} \cor \neglit{\cst{d}}\}$.
%
Either clause is subsumed by an active clause, so the prover deletes it.
It stops with the active set
$\{\cst{c} \cor \cst{d}{,}\; \cst{b} \cor \neglit{\cst{d}}\}$, which is
unsaturated because $\iota$ is no longer redundant. The invariant is broken.
\end{exa}

\begin{exa}
\label{ex:cex-predelim-unsat}
In Example~\ref{ex:cex-predelim-sat}, the initial clause set was satisfiable.
If it is unsatisfiable, we can even lose refutational
completeness. To see why, we add the unit clauses
$\neglit{\cst{b}}$ and $\neglit{\cst{c}}$ to the initial clause set of
Example~\ref{ex:cex-predelim-sat} to make it unsatisfiable. We repeat the same
steps as above, including the subsumptions at the end,
yielding the passive set $\{\MAX{\neglit{\cst{b}}}, \MAX{\neglit{\cst{c}}}\}$ and
the active set $\{\cst{c} \cor \MAX{\cst{d}}{,}\; \cst{b} \cor \MAX{\neglit{\cst{d}}}\}$. Then,
making $\neglit{\cst{b}}$ and $\neglit{\cst{c}}$ active triggers no
inferences. The prover stops with an
unsatisfiable four-clause active set that does not contain the empty clause.
\end{exa}

A solution could be to move all active clauses to the passive set at step~$k$
or later, but this would be costly, since it would force the prover to redo
inferences whose conclusions might then have to be simplified or subsumed again.
Instead, we salvage the existing completeness proof for $\mathsf{GC}$, by
resolving the issues concerning splitting and the $\mathsf{GC}$ invariant. Our
approach is to weaken the redundancy criterion slightly, enough both to disable
splitting on $\mathbf{P}$-predicates and to ensure that
inferences such as $\iota$ in Examples
\ref{ex:cex-predelim-sat}~and~\ref{ex:cex-predelim-unsat} are performed. The
required weakening is so mild that it does not invalidate any practical
simplification or subsumption techniques we are aware of, except of course
splitting.

In accordance with the saturation framework, let $\mathbf{F}$ be the set of
first-order $\mupSigma$-clauses, let $\mathbf{G}$ be its ground subset,
and let $\gnd$ be a function that maps an $\mathbf{F}$-clause to the set of its
$\mathbf{G}$-clause instances and analogously for $\mathbf{F}$-inferences. We
define an extension $\mathbf{G}\Wksup$ of $\mathbf{G}$ for
$\mupSigma\Wksup$-clauses in an ad hoc nonclassical logic reminiscent
of paraconsistent logics \cite{carnielli-et-al-2007}. The objective is to
disallow the entailment that makes splitting and Examples
\ref{ex:cex-predelim-sat}~and~\ref{ex:cex-predelim-unsat} possible.
%
The signature~$\mupSigma\Wksup$ extends $\mupSigma$ with a
distinguished predicate symbol $\false$ that is interpreted differently
from~$\top$. For $\mupSigma\Wksup$, the Boolean type $\prop$ may be
interpreted as any set of cardinality at least 2.

\begin{defi}
The operator $\Wk{}$ translates $\mupSigma$-literals to
$\mupSigma\Wksup$-literals as follows, where $\cst{p} \in \mathbf{P}$,
$\cst{q} \notin \mathbf{P}$, and $s, t$ are non-Boolean terms:
%
%\begin{linenomath*}
\begin{align*}
  \Wk{\cst{p}(\tuple{t}{})} & = \cst{p}(\tuple{t}{}) \noteq \false
& \Wk{\cst{q}(\tuple{t}{})} & = \cst{q}(\tuple{t}{}) \eq \true
& \Wk{(s \eq t)} & = s \eq t \\[-\jot]
  \Wk{\lnot\,\cst{p}(\tuple{t}{})} & = \cst{p}(\tuple{t}{}) \noteq \true
& \Wk{\lnot\,\cst{q}(\tuple{t}{})} & = \cst{q}(\tuple{t}{}) \noteq \true
& \Wk{(s \noteq t)} & = s \noteq t
\end{align*}
%\end{linenomath*}
%
The operator is lifted elementwise to $\mathbf{G}$-clauses and
$\mathbf{G}$-clause sets.
The \emph{weak entailment} $\modelsWk$ over $\mathbf{G}$-clause sets
is defined via an encoding into $\mupSigma\Wksup$-clauses:
$M \modelsWk N$ if and only if $\Wk{M} \models \Wk{N}$.
The lifting to $\mathbf{F}$-clauses and $\mathbf{F}$-clause set is achieved
in the standard way via grounding.
\end{defi}

The following property of weak entailment will allow us to eliminate
$\mathbf{P}$-predicates without losing completeness:

\begin{lemma}
\label{lem:p-Wk-entail}
Let $C$ be a clause that contains the predicate symbol $\cst{p} \in
\mathbf{P}$ and $D$ be a clause that does not contain $\cst{p}$. If $N
\cup \{C\} \modelsWk \{D\}$, then $N \modelsWk \{D\}$.
\end{lemma}

\begin{proof}
Suppose $\mathscr{J} \models \Wk{N}$. We will define $\mathscr{J}'$ so that
$\mathscr{J}' \models \Wk{N} \cup \{\Wk{C}\}$, retrieve $\mathscr{J}'
\models \Wk{D}$, and then argue that $\mathscr{J} \models \Wk{D}$. We
take $\mathscr{J}'$ to coincide with $\mathscr{J}$ except that we extend the
domain for $\prop$ with one fresh value and use this value as the
interpretation of $\cst{p}(\tuple{t}{})$ for all argument tuples $\tuple{t}{}$.
This modification makes any $\cst{p}$ literal of $\Wk{C}$ true, and it
preserves the truth of $\Wk{N}$. By the hypothesis, $\mathscr{J}' \models
\Wk{D}$. And since $\cst{p}$ does not occur in $D$, we have $\mathscr{J}
\models \Wk{D}$.\qed
\end{proof}

Note that the above lemma does not hold for classical entailment $\models$;
indeed, $\{\cst{p} \cor \cst{q}{,}\; \neglit{\cst{p}} \cor \cst{q}\} \models
\{\cst{q}\}$. On the other hand, the law of excluded middle does hold for weak
entailment: $\modelsWk \cst{p} \cor \lnot\,\cst{p}$. In fact, all classical
clausal tautologies are tautologies for $\modelsWk$.

The standard redundancy criterion $\Red$ is obtained by lifting a criterion on
$\mathbf{G}$-clauses to $\mathbf{F}$-clauses. The same construction can be
replicated using $\modelsWk$ instead of $\models$, yielding the \emph{weak
redundancy criterion} $\RedWk$.
It is easy to check that the usual simplification techniques implemented in
superposition provers can be justified using $\RedWk$. Specifically,
this concerns the following rules described by Schulz \cite[Sects.\
2.3.1~and~2.3.2]{ss-02-brainiac}:\
deletion of duplicated literals, 
deletion of resolved literals, 
syntactic tautology deletion,
semantic tautology deletion,
rewriting of negative literals,
positive simplify-reflect,
negative simplify-reflect,
clause subsumption, and
equality subsumption.
Moreover, rewriting of positive literals is possible if the
rewriting clause is smaller than the rewritten
clause (a condition that is also needed with $\models$ but omitted by Schulz).
Finally, destructive equality resolution cannot be justified with $\models$,
let alone $\modelsWk$.

We instantiate the saturation framework with $(\FInf,\RedWk)$ to obtain a
given clause prover \textsf{GC}. The prover operates on sets of labeled clauses
$(C, l)$, where $C$ is a standard clause and $l \in \Lab$ is a label.
The $\mathsf{active}$ label identifies active clauses; all other clauses are
passive. The prover takes the form of two rules,
\textsc{Process} and \textsc{Infer}, restricted to prevent the introduction of
$\mathbf{P}$-predicates. We extend it with a third rule,
\textsc{PredElim}, for predicate elimination, and call the extended prover
\textsf{GCP}. The rules are as follows, using again the framework notations:

\begin{description}[labelwidth=\widthof{\rm\textsc{PredElim}\,}]
\itemsep1\jot
\item[\rm\textsc{Process}\,]
  $\mathcalx{N} \cup \mathcalx{M}
   \,\Longrightarrow_\mathsf{GCP}\,
   \mathcalx{N} \cup \mathcalx{M}'$
\\
  where $\mathcalx{M} \subseteq \LRedFWksub(\mathcalx{N} \cup \mathcalx{M}')$,
  $\mathcalx{M}'{\downarrow}_\mathsf{active} = \emptyset$,
  and \\
  $\mathbf{P}(\lfloor\mathcalx{M}'\rfloor) \subseteq \mathbf{P}(\lfloor\mathcalx{N} \cup \mathcalx{M}\rfloor)$

\item[\rm\textsc{Infer}\,]
  $\mathcalx{N} \cup \{(C, l)\}
   \,\Longrightarrow_\mathsf{GCP}\,
   \mathcalx{N} \cup \{(C, \ensuremath{\mathsf{active}})\} \cup \mathcalx{M}$
\\
  where $l \not= \ensuremath{\mathsf{active}}$,
  $\mathcalx{M}{\downarrow}_\mathsf{active} = \emptyset$, \\
  $\FInf(\lfloor\mathcalx{N}{\downarrow}_\mathsf{active}\rfloor, \{C\}) \subseteq
    \RedIWk^{{\cap}\gnd}(\lfloor\mathcalx{N}\rfloor \cup \{C\} \cup \lfloor\mathcalx{M}\rfloor)$,
  and \\
  $\mathbf{P}(\lfloor\mathcalx{M}\rfloor) \subseteq \mathbf{P}(\lfloor\mathcalx{N}\rfloor \cup \{C\})$

\item[\rm\textsc{PredElim}\,]
  $\mathcalx{N} \cup \mathcalx{M}
   \,\Longrightarrow_\mathsf{GCP}\,
   \mathcalx{N} \cup \mathcalx{M}'$
\\
  where $\mathcalx{N} \cup \mathcalx{M} \rhd_\mathbf{P} \mathcalx{N} \cup \mathcalx{M}'$
  and $\mathcalx{M}'{\downarrow}_\mathsf{active} = \emptyset$
\end{description}

Here is a summary of the main framework notations:
%
\begin{itemize}
\item The letters $\mathcalx{M}, \mathcalx{N}$ range over sets of labeled clauses.
  $\mathcalx{M}{\downarrow}_l$ denotes the subset of clauses in $\mathcalx{M}$
  labeled with $l$. The operator $\lfloor\phantom{\cdot}\rfloor$ erases all
  labels in a labeled clause or clause set.

\smallskip

\item $\FInf(N)$ denotes the set of all base calculus inferences with premises
  in $N$, and $\FInf(N, M) = \FInf(N \cup M) \setminus \FInf(N \setminus M)$.
  The same notations are also available for the straightforward extension $\FLInf$
  of $\FInf$ with labels.

\smallskip

\item $\LRedWksub$ is the extension of the standard redundancy
  criterion $\RedWk$ defined using $\modelsWk$ to nonground labeled clauses
  with subsumption ($\sqsubset$).

\smallskip

\item Given a sequence $(\mathcalx{N}{}_i)_i$, its \emph{limit} (\emph{inferior})
  is $\mathcalx{N}{}_\infty = \bigcup_i \bigcap_{j\ge i} \mathcalx{N}{}_j$.
\end{itemize}

\begin{sloppypar}
   The completeness proof follows the invariance-based argument found in the
   article by Waldmann et al \cite{waldmann-et-al-202x-article}.
\end{sloppypar}

\begin{lemma}
\label{lem:gcp-derivations-are-red-black-derivations}
Every $\Longrightarrow_\mathsf{GCP}$-derivation is a mixed derivation.
\end{lemma}

\begin{proof}
The cases for \textsc{Process} and \textsc{Infer} are almost as in Waldmann et
al., with adjustments to show that $\cst{P}$-predicates cannot reappear once
they have disappeared. The case for \textsc{ElimPred} is trivial.
\end{proof}

Let $\Inv\Wksup_\mathcalx{N}(k)$ denote the condition
$\textstyle\FLInf(\mathcalx{N}{}_k{\downarrow}_\mathsf{active}) \subseteq
\LRedIWk(\mathcalx{N}{}_k)$.
Notice the difference with the definition of the key invariant
$\Inv_\mathcalx{N}$ in the saturation framework, whose right-hand side is
$\bigcup_{i=0}^k \RedI^{\Lab}(\mathcalx{N}{}_i)$. We cannot use the
big union ${\bigcup}$ starting at $i = 0$ because we will need to truncate a
sequence prefix of an a priori unknown length. The argument will still work
thanks to monotonicity properties of redundancy criteria.

\begin{lemma}
\label{lem:gcp-invar-gcp}
Let $(\mathcalx{N}{}_i)_i$ be a $\Longrightarrow_\mathsf{GCP}$-derivation.
If
$\mathcalx{N}{}_0{\downarrow}_\mathsf{active} = \emptyset$,
then
$\Inv\Wksup_\mathcalx{N}(k)$ holds for all indices~$k$.
\end{lemma}

\begin{proof}
The base case is as in Waldmann et al.

For \textsc{Process} and \textsc{Infer}, the proof is essentially as in Waldmann
et al., except that we also need to show that
$\LRedIWk(\mathcalx{N}{}_k) \subseteq \LRedIWk(\mathcalx{N}{}_{k+1})$.
This is a consequence of
$\mathcalx{N}{}_k \rhd_{\LRedWksub} \mathcalx{N}{}_{k+1}$ and of
properties (R2) and (R3) of redundancy criteria.

\begin{sloppypar}

   A new case to consider is that of a \textsc{PredElim} transition
   $\mathcalx{N}{}_{k} \Longrightarrow_\mathsf{GCP} \mathcalx{N}{}_{k+1}$.
   Let $\mathcalx{N}{}_{k} = \mathcalx{N} \cup \mathcalx{M}
   \Longrightarrow_\mathsf{GCP} \mathcalx{N} \cup \mathcalx{M}' = \mathcalx{N}{}_{k+1}$,
   where $\mathcalx{N} \cup \mathcalx{M} \rhd_\mathbf{P}
   \mathcalx{N} \cap \mathcalx{M}'$ and
   $\mathcalx{M}'{\downarrow}_\mathsf{active} = \emptyset$.
   We assume without loss of generality that $\mathcalx{M} \cap \mathcalx{M}' =
   \emptyset$.
   Let $\cst{p}$ be the eliminated predicate. Note that $\cst{p}$ occurs in every
   clause in $\mathcalx{M}$ but in none of the clauses in $\mathcalx{N}$ or
   $\mathcalx{M}'$.
   We must show $\FLInf(\mathcalx{N}{}_{k+1} {\downarrow}_\mathsf{active})
   \subseteq \LRedIWk(\mathcalx{N}{}_{k+1})$.
   As for \textsc{Process}, we have the inclusion
   $\FLInf(\mathcalx{N}{}_{k+1} {\downarrow}_\mathsf{active})
   \subseteq \FLInf(\mathcalx{N}{}_k {\downarrow}_\mathsf{active})$,
   by the side condition that
   $\mathcalx{M}'{\downarrow}_\mathsf{active} = \emptyset$.
   Moreover, by the induction hypothesis,
   $\FLInf(\mathcalx{N}{}_k {\downarrow}_\mathsf{active})
   \subseteq \LRedIWk(\mathcalx{N}{}_k)$. Thus,
   $\FLInf(\mathcalx{N}{}_k {\downarrow}_\mathsf{active})
   \subseteq \LRedIWk(\mathcalx{N} \cup \mathcalx{M})$.
\end{sloppypar}

Let $\iota \in \FLInf(\mathcalx{N} {\downarrow}_\mathsf{active})$.
By the argument above, we have
$\iota \in \LRedIWk(\mathcalx{N} \cup \mathcalx{M})$.
We must
show $\iota \in \LRedIWk(\mathcalx{N} \cup \mathcalx{M}')$.
%
By definition of $\LRedIWk$, it suffices to show that for
every ground instance $(C_n,\ldots,C_1,C_0)$ of $\iota$, there exists a finite
clause set $\mathcalx{D} \subseteq \gnd(\mathcalx{N}) \cup \gnd(\mathcalx{M}')$
that is $\prec$-smaller than $C_1$ and such that
$\{C_n,\ldots,C_2\} \cup \mathcalx{D} \modelsWk \{C_0\}$.
Without loss of generality, we assume that $\mathcalx{D}$ is the smallest such
set.

By definition of $\mathcalx{N}$, $\cst{p}$ cannot occur in $C_0$.
By Lemma~\ref{lem:p-Wk-entail}, if $\cst{p}$ occurred in $D \in \mathcalx{D}$,
we could remove $D$, but this would mean $\mathcalx{D}$ is not minimal.
As a result, $\mathcalx{D}$ cannot contain clauses from $\gnd(\mathcalx{M})$ and
hence $\mathcalx{D} \subseteq \gnd(\mathcalx{N})$. Thus,
$\iota \in \LRedIWk(\mathcalx{N})$. By property (R2) of
redundancy criteria, we have the desired result: $\iota \in
\LRedIWk(\mathcalx{N} \cup \mathcalx{M}')$.
\end{proof}

\begin{lemma}
\label{lem:gcp-invar-infinity}
Let $(\mathcalx{N}{}_i)_i$ be a
$\rhd_{\LRedWksub}$-derivation.
If $\Inv\Wksup_\mathcalx{N}(i)$ holds for all indices~$i$, then
$\FLInf(\mathcalx{N}{}_\infty{\downarrow}_\mathsf{active}) \subseteq
\bigcup_i \LRedIWk(\mathcalx{N}{}_i)$
holds.
\end{lemma}

\begin{proof}
We assume $\iota \in \FLInf(\mathcalx{N}{}_\infty {\downarrow}_\mathsf{active})$
and show $\iota \in \bigcup_i \LRedIWk(\mathcalx{N}{}_i)$
for some arbitrary $\iota$.
%
For $\iota$ to belong to $\FLInf(\mathcalx{N}{}_\infty {\downarrow}_\mathsf{active})$, all of
its finitely many premises must be in $\mathcalx{N}{}_\infty {\downarrow}_\mathsf{active}$.
Therefore, there must exist an index $k$ such that
$\mathcalx{N}{}_k {\downarrow}_\mathsf{active}$ contains all of them, and therefore
$\iota \in \FLInf(\mathcalx{N}{}_k {\downarrow}_\mathsf{active})$. Since $\Inv_\mathcalx{N}(k)$
holds, $\iota \in \LRedIWk(\mathcalx{N}{}_k)$.
Hence, $\iota \in \bigcup_i \LRedIWk(\mathcalx{N}{}_i)$.
\end{proof}

\begin{lemma}
\label{lem:fair-gcp-derivations}
Let $(\mathcalx{N}{}_i)_i$ be a $\Longrightarrow_\mathsf{GCP}$-derivation.
If
$\mathcalx{N}{}_0{\downarrow}_\mathsf{active} = \emptyset$
and
$\mathcalx{N}{}_\infty{\downarrow}_l = \emptyset$ for every label
$l \not= \ensuremath{\mathsf{active}}$,
then there exists an index $k$ such that
$(\mathcalx{N}{}_{k+i})_i$ is a fair $\rhd_{\LRedWksub}$-derivation.
\end{lemma}

\begin{sloppypar}
\begin{proof}
By Lemma~\ref{lem:gcp-derivations-are-red-black-derivations}, there must exist
an index $k$ such that the sequence $(\mathcalx{N}{}_{k+i})_i$ is a pure
$\rhd_{\LRedWksub}$-derivation.
By Lemma~\ref{lem:gcp-invar-gcp},
$\Inv\Wksup_\mathcalx{N}(k+i)$ holds for all indices~$i$.
Hence, by Lemma~\ref{lem:gcp-invar-infinity},
$\FLInf(\mathcalx{N}{}_\infty{\downarrow}_\mathsf{active}) \subseteq
\bigcup_{i} \LRedIWk(\mathcalx{N}{}_{k+i})$.
By the second hypothesis, this inclusion simplifies to
$\FLInf(\mathcalx{N}{}_\infty) \subseteq
\bigcup_{i} \LRedIWk(\mathcalx{N}{}_{k+i})$.
\end{proof}
\end{sloppypar}

\begin{theorem}
\label{thm:gcp-complete}
Let $(\mathcalx{N}{}_i)_i$ be a $\Longrightarrow_\mathsf{GCP}$-derivation
with
$\mathcalx{N}{}_0{\downarrow}_\mathsf{active} = \emptyset$
and
$\mathcalx{N}{}_\infty{\downarrow}_l = \emptyset$ for every label
$l \not= \ensuremath{\mathsf{active}}$.
If $\lfloor \mathcalx{N}{}_0\rfloor$ is unsatisfiable,
then
some $\mathcalx{N}{}_i$
contains the empty clause with some arbitrary label.
\end{theorem}

\begin{proof}
By Lemma~\ref{lem:fair-gcp-derivations},
we know that there exists an index $k$ such that
$(\mathcalx{N}{}_{k+i})_i$ is a fair $\rhd_{\LRedWksub}$-derivation.
Moreover, since
$\rhd_{\LRedWksub}$ and
$\rhd_\mathbf{P}$ preserve unsatisfiability
(by (R1) of redundancy criteria, Khasidashvili and Korovin's Theorem~1, and our
Theorem~\ref{thm:pes-sat-equiv}), we have that
$\lfloor \mathcalx{N}{}_k\rfloor$ is unsatisfiable.
Since the base calculus $(\FLInf,\allowbreak\LRed)$ is assumed to be statically
refutationally complete with respect to $\models$, the calculus
$(\FLInf,\allowbreak\LRedWk)$ with a weaker redundancy criterion is also
statically complete with respect to $\models$, and by the saturation framework,
$(\FLInf,\allowbreak\LRedWksub)$ preserves this. Exploiting the equivalence of
static and dynamic completeness, we conclude that some $\mathcalx{N}{}_{k+i}$ must
contain a labeled empty clause.
\end{proof}
\end{rep}
